Gammapy: A Python package for gamma-ray astronomy

### 2022-12-06 src/ms-review.tex

The DL3 level data represented by the \code{Observation} class consist
of two types of elements: first a list of \gammaray events with relevant physical
quantities such as estimated energy, direction and arrival
times, which is represented by the \code{EventList} class/. Second a set of
associated IRFs, providing the response of the system, typically
factorised in independent components as described in
Section~\ref{ssec:gammapy-irf}. The separate handling of event lists and IRFs
additionally allows for data from non-IACT \gammaray instruments to be read. For
example, to read \fermi data, the user can read separately their event list
(already compliant with the \gadf specifications) and then find the appropriate
IRF class representing the response functions provided by \fermi, see
example in Section~\ref{ssec:multi-instrument-analysis}.
%
\begin{figure}[ht!]
	\import{code-examples/generated/}{gp_data}
	\caption{
        Using \code{gammapy.data} to access DL3 level data with a \code{DataStore} object.
        Individual observations can be accessed by their unique integer observation id number.
        The actual events and instrument response functions can be accessed
        as attributes on the \code{Observation} object, such as \code{.events}
        or \code{.aeff} for the effective area information.
    }
	\label{fig*:minted:gp_data}
\end{figure}
%

\subsection{gammapy.irf}
\label{ssec:gammapy-irf}
%
%\begin{figure}
%	%	\import{code-examples/generated/}{gp_makers}
%
%	\caption{In Fig.~\ref{ig*:minted:irf_examples}, we show some example of Aeff,
%		PSF, Edisp and BKG read and plotted from a typical \irf.}
%	\label{ig*:minted:irf_examples}
%\end{figure}
%
The \code{gammapy.irf} sub-package contains all classes and functionality
to handle IRFs in a variety of formats.
Usually, \irfs store instrument properties in the form of multi-dimensional
tables, with quantities expressed in terms of energy (true or reconstructed),
off-axis angles or cartesian detector coordinates. The main information stored in
the common \gammaray \irfs are the effective area (Aeff), energy dispersion
(Edisp), point spread function (PSF) and background rate (BKG). The \code{gammapy.irf}
sub-package can open and access specific \irf extensions,
interpolate and evaluate the quantities of interest on both energy and spatial
axes, convert their format or units in different kinds, plot or write them into
output files. In the following, we list the main classes of the
sub-package:

\subsubsection{Effective Area}
\gammapy provides the class \code{EffectiveAreaTable2D} to
manage Aeff, which is usually defined in terms of true energy and offset angle.
The class functionalities offer the possibility to read from files or to create
it from scratch. {\code{EffectiveAreaTable2D} can also convert, interpolate,
write, and evaluate the effective area for a given energy and offset angles, or
even plot the multi-dimensional Aeff table.

\subsubsection{Point Spread Function}
\gammapy allows user to treat different kinds of PSFs,
in particular, parametric multi-dimensional Gaussians (\code{EnergyDependentMultiGaussPSF})
or King profile functions (\code{PSFKing})
one. The \code{EnergyDependentMultiGaussPSF} class is able to handle up to three
Gaussians, defined in terms of amplitudes and sigma given for each true energy
and offset angle bin. Similarly, \code{PSFKing takes} into account the gamma and
sigma parameters. The general \code{ParametricPSF} class allows users to create a
custom PSF with a parametric representation different from Gaussian(s) or King profile(s).
The generic \code{PSF3D} class stores a radial symmetric profile of a
PSF to represent non-parametric shapes, againa depending on true energy 
of offset form the poiting position.

@@

\subsubsection{Energy Dispersion}
for \iact, it is generally given in terms of the so-called
 migration parameter ($\mu$), which is defined as the ratio between the
reconstructed energy and the true energy. This ratio should be as close as one
and its dispersion can assume the shape of a Gaussian (or even more complex
distributions). Migration parameter is given at each offset angle and
reconstructed energy. The main sub-classes are \code{EnergyDispersion2D},
designed to interpret Edisp, \code{EDispKernelMap}, which builds an Edisp kernel
map, i.e., a 4-dimensional WCS map where at each sky-position is associated an
Edisp kernel. The latter is a representation of the Edisp as a function of the
true energy only thanks to the sub-class \code{EDispKernel}.

\subsubsection{Background model}
The BKG can be represented in \gammapy as either 1) a 2D map (\code{Background2D})
of count rate normalised per steradians and energy at different
reconstructed energies and offset angles or 2) as rate per steradians and energy, as a
function of reconstructed energy and detector coordinates (\code{Background3D}).
In the former, the background is expected to follow a radially symmetric shape,
while in the latter, it can be more complex.

@@

The models follow a naming scheme which contains the category as a suffix to
the class name. The spectral models include a special class of normed models,
named using the \code{NormSpectralModel} suffix.
These spectral models feature a dimension-less \textit{norm} parameter
instead of an \text{amplitude} parameter with physical units. They
must be used along with another spectral model, as a multiplicative correction
factor according to their spectral shape. They can be typically used for
adjusting template based models, or adding taking into account
the absorbtion effect on 
\gammaray spectra caused by the extra-galactic background light (EBL) (\code{EBLAbsorptionNormSpectralModel})
model. \gammapy supports a variety of EBL absorbtion models, such as

\cite{Franceschini2008}, \cite{Finke2010} and \cite{Dominguez2011}.

@@

The model gallery provides a visual overview of the available models in
\gammapy. Most of the analytic models  commonly used in \gammaray astronomy are
built-in. We also offer a wrapper to radiative models implemented in the Naima
package~\cite{naima}. The modeling framework can be easily extended with
user-defined models. For example agnpy models that describe leptonic radiative
processes in jetted Active Galactic Nuclei (AGN) can be wrapped into
\gammapy~\citep[see Section 3.5 of ][]{2021arXiv211214573N} .

@@

The Fit class provides a uniform interface to multiple fitting backends:
\iminuit~\citep{iminuit}, “scipy”,~\citep{2020SciPy-NMeth}, and
“sherpa”~\citep{sherpa-2005,sherpa-2011}. Note that, for now, covariance matrix
and errors are computed only for the fitting with \iminuit. However depending on
the problem other optimizers can better perform, so sometimes it can be useful
to run a pre-fit with alternative optimization methods. In future we plan to
extend the supported fitting backends, including for example solutions based on Markov chain Monte Carlo.
\footnote{a prototype is available in gammapy-recipes,
	\url{https://gammapy.github.io/gammapy-recipes/_build/html/notebooks/mcmc-sampling-emcee/mcmc_sampling.html}
}

@@

Optionally it can also compute more advanced quantities
such as assymmetric flux errors, flux upper limits
and one dimensional profiles of the fit statistic,
which show how the likelihood functions varies with
the flux norm parameter around the fit minimum.
This information is useful for inspecting the quality
of a fit, for which asymptomatically a parabolic
shape of the profile is expected at the best fit
values (\todo{Reference to modeling / fitting section?}).

@@

Code example ~\ref{fig*:minted:gp_estimators} shows how to use
the \code{TSMapEstimator} objects with a given input \code{MapDataset}.
In addition to the model it allows to specify the energy
bins of the resulting flux and TS maps.

@@

\subsection{gammapy.analysis}
\label{ssec:gammapy-analysis}
The \code{gammapy.analysis} sub-module provides a high-level interface (HLI) for the most
common use cases identified in \gammaray analyses. The included classes and methods
 can be used in Python scripts, notebooks or as commands within IPython
sessions. The HLI can also be used to automatise
workflows driven by parameters declared in a configuration file in YAML format.
This way, a full analysis can be executed via a single command line taking the
configuration file as input.

The \code{Analysis} class has the responsibility of orchestrating of the workflow
defined in the configuration \code{AnalysisConfig} objects and triggering the execution of
the \code{AnalysisStep} classes that define the identified common use cases. These
steps include the following: observations selection with the \code{DataStore},  data
reduction, excess map computation, model fitting, flux points estimation, and
light curves production. The structure of the \code{Analysis} class allows to define
custom AnalysisStep, so one could design and configure arbitrarily complex
workflows. \todo{the latter is not true for v1.0..., include example YAML file?}

@@

\subsection{1D Analysis}
\label{ssec:1d-analysis}
One of the most common analysis cases in \gammaray astronomy is measuring the
spectrum of a source in a given region defined on the sky, in conventional
astronomy also called \textit{aperture photometry}. The spectrum is typically measured
in two steps: first a parametric spectral model is fitted to the data and
secondly flux points are computed in a pre-defined set of energy bins. The
result of such an analysis performed on three simulated CTA observations is
shown in Fig.~\ref{fig:cta_galactic_center}. In this case the spectrum was
measured in a circular aperture centered on the Galactic Center, in
\gammaray~astronomy often called "on region". For such analysis the users first
chooses a region of interest and energy binning, both defined by a
\code{RegionGeom}. In a second step the events and the IRFs are binned
into maps of this geometry, by the \code{SpectrumDatasetMaker}. All the data and
reduced IRFs are bundled into a \code{SpectrumDataset}. To estimate
the expected background in the ON region a 'reflected regions' background
method was used~\cite{Berge07}, represented in \gammapy by the
\code{ReflectedRegionsBackgroundMaker} class. The resulting reflected regions are
illustrated for all three observations on top of the map of counts. After
reduction of the data it was modelled using a forward-folding method and a
power law spectral shape, using the \code{PowerLawSpectralModel} and \code{Fit} class.
Based on this best fit model the final flux points and corresponding
log-likelihood profiles are computed using the \code{FluxPointsEstimator}.

@@

The result of the joint fit is displayed in
Fig.~\ref{fig:multi_instrument_analysis}. We remark that the objective of this
exercise is illustrative. We display the flexibility of \gammapy in
simultaneously fitting multi-instrument data even at different levels of
reduction, without aiming to provide a new measurement of the Crab Nebula
spectrum.

@@

\subsection{Broadband SED Modeling}
\label{ssec:broadband-sed-modeling}
By combining \gammapy with astrophysical modelling codes, users can also fit
astrophysical spectral models to \gammaray data. In \gammaray
astronomy one typically observes two radiation production
mechanisms, the so-called hadronic and leptonic scenarios.
There are several Python packages that are able to model
the \gammaray emission, given a physical scenario. Among those
packages are  Agnpy \citep{agnpy}, Naima \citep{naima}, Jetset \citep{jetset}
and Gamera \citep{gamera}.
Tyically those emission models predict broadband emission from
radio, up to the very high-energy \gammaray range.
By relying on the multiple dataset types in \gammapy those
data can be combined to constrain such a broadband emission model.
\gammapy provides a built-in \code{NaimaSpectralModel} that allows
users to wrap a given astrophysical emission model from the
naima package and fit it directly to \gammaray data.

As an example of this application, we use the same multi-instrument
dataset described in the previous section and we fit it with an inverse
Compton model computed with naima and wrapped in the \gammapy fitting
 routine with the NaimaSpectraModel class. We consider a log-parabolic
electron distribution scattering: the synchrotron radiation produced by
the same electrons; near and far infrared photon fields and the cosmic
microwave background (CMB). We adopted the prescription on the target
photon fields provided in the
documentation of the \textit{Naima} package \url{https://naima.readthedocs.io/en/latest/examples.html#crab-nebula-ssc-model}.
The best-fit inverse Compton spectrum is represented with a red dashed line in
Fig.~\ref{fig:multi_instrument_analysis}.

@@

Several automated tasks are set as \github
actions\footnote{\url{https://github.com/features/actions}}, blocking the
processes and alerting developers when fails occur. This is the case of the
continuous integration workflow, which monitors the execution of the test coverage
suite\footnote{\url{https://pytest.org}} using datasets from the
\textit{gammapy-data} repository. Tests scan not only the codebase but also the
code snippets present in docstrings of the scripts and in the RST documentation
files, as well as in the tutorials provided in the form of Jupyter notebooks.

Other automated tasks, executing in the \textit{gammapy-benchmarks} repository,
are responsible for numerical validation tests and benchmarks monitoring. Also
tasks related with the release process are partially automated, and every
contribution to the codebase repository triggers the documentation building and
publishing workflow within the \textit{gammapy-docs} repository
(see Sec.~\ref{ssec:software-distribution} and Sec.~\ref{ssec:documentation-and-user-support}).

@@

We have also set up a solution for users to share recipes as Jupyter notebooks
that do not fit in the \gammapy core documentation but which may be relevant as
specific use cases. Contributions happen via pull requests to the
\textit{gammapy-recipes} \github repository and merged after a short review. All
notebooks in the repository are tested and published in the \gammapy recipes
webpage\footnote{\url{https://gammapy.\github.io/gammapy-recipes}} automatically
using \github actions.

A growing community of users is gathering around the Slack
messaging\footnote{\url{https://gammapy.slack.com}} and \github
discussions\footnote{\url{https://\github.com/gammapy/gammapy/discussions}}
support forums, providing valuable feedback on the \gammapy functionalities,
interface and documentation. Other communication channels have been set like
mailing lists, a Twitter account\footnote{\url{https://twitter.com/gammapyST}},
regular public coding sprint meetings, hands-on session within collaborations,
weekly development meetings, etc.

@@

\section{Reproducibility}
\label{sec:reproducibility}
One of the most important goals of the \gammapy project is to support open and
reproducible science results. Thus we decided to write this manuscript
openly and publish the Latex source code along with the associated
Python scripts to create the figures
in \url{https://github.com/gammapy/gammapy-v1.0-paper}.
This \github repository also documents the history of the creation
and evolution of the manuscript with time. To simplify the reproducibility
of this manuscript including figures and text, we relied on the tool
\textit{showyourwork}~\citep{Luger2021}. This tool coordinates the building
process and software as well as data dependencies such, that the complete
manuscript can be reproduced with a single \code{make} command, after
downloading the source repository. See \url{https://github.com/gammapy/gammapy-v1.0-paper/blob/main/README.md} for detailed
instructions. Almost all figures in this manuscript provide a link
to a Python script, that was used to produce it. This means all
example analyses presented in Sec.\ref{sec:applications} link to
actually working Python source code.

>>>>

The DL3 level data represented by the \code{Observation} class consist
of two types of elements: first a list of \gammaray events with relevant physical
quantities such as estimated energy, direction and arrival
times, which is represented by the \code{EventList} class. Second a set of
associated IRFs, providing the response of the system, typically
factorised in independent components as described in
Section~\ref{ssec:gammapy-irf}. The separate handling of event lists and IRFs
additionally allows for data from non-IACT \gammaray instruments to be read. For
example, to read \fermi data, the user can read separately their event list
(already compliant with the \gadf specifications) and then find the appropriate
IRF class representing the response functions provided by \fermi, see
example in Section~\ref{ssec:multi-instrument-analysis}.
%
\begin{figure}[ht!]
	\import{code-examples/generated/}{gp_data}
	\caption{
        Using \code{gammapy.data} to access DL3 level data with a \code{DataStore} object.
        Individual observations can be accessed by their unique integer observation id number.
        The actual events and instrument response functions can be accessed
        as attributes on the \code{Observation} object, such as \code{.events}
        or \code{.aeff} for the effective area information.
    }
	\label{fig*:minted:gp_data}
\end{figure}
%

\subsection{gammapy.irf}
\label{ssec:gammapy-irf}
%
%\begin{figure}
%	%	\import{code-examples/generated/}{gp_makers}
%
%	\caption{In Figure~\ref{ig*:minted:irf_examples}, we show some example of Aeff,
%		PSF, Edisp and BKG read and plotted from a typical \irf.}
%	\label{ig*:minted:irf_examples}
%\end{figure}
%
The \code{gammapy.irf} sub-package contains all classes and functionality
to handle IRFs in a variety of formats.
Usually, \irfs store instrument properties in the form of multi-dimensional
tables, with quantities expressed in terms of energy (true or reconstructed),
off-axis angles or cartesian detector coordinates. The main information stored in
the common \gammaray \irfs are the effective area, energy dispersion,
point spread function and background rate. The \code{gammapy.irf}
sub-package can open and access specific \irf extensions,
interpolate and evaluate the quantities of interest on both energy and spatial
axes, convert their format or units in different kinds, plot or write them into
output files. In the following, we list the main classes of the
sub-package:

\subsubsection{Effective Area}
\gammapy provides the class \code{EffectiveAreaTable2D} to
manage the effective area, which is usually defined in terms of true energy and offset angle.
The class functionalities offer the possibility to read from files or to create
it from scratch. {The \code{EffectiveAreaTable2D} class can also convert, interpolate,
write, and evaluate the effective area for a given energy and offset angles, or
even plot the multi-dimensional effective area table.

\subsubsection{Point Spread Function}
\gammapy allows user to treat different kinds of PSFs,
in particular, parametric multi-dimensional Gaussians (\code{EnergyDependentMultiGaussPSF})
or King profile functions (\code{PSFKing}). The \code{EnergyDependentMultiGaussPSF}
 class is able to handle up to three
Gaussians, defined in terms of amplitudes and sigma given for each true energy
and offset angle bin. Similarly, \code{PSFKing} takes into account the gamma and
sigma parameters. The general \code{ParametricPSF} class allows users to create a
custom PSF with a parametric representation different from Gaussian(s) or King profile(s).
The generic \code{PSF3D} class stores a radial symmetric profile of a
PSF to represent non-parametric shapes, againa depending on true energy 
of offset form the poiting position.

@@

\subsubsection{Energy Dispersion}
For \iacts, the energy dispersion is typically parameterised in terms of the so-called
migration parameter ($\mu$), which is defined as the ratio between the
reconstructed energy and the true energy. By definition the mean of this ratio is
close to unity for a small energy bias and its distribution can 
be typically described by a Gaussian, however more complex
shapes are also common. The migration parameter is given at each offset angle and
reconstructed energy. The main sub-classes are the \code{EnergyDispersion2D} which is
designed to handle the raw instrument description. Then there is the \code{EDispKernelMap},
which contains an energy disperion matrix per sky position. I.e., a 4-dimensional WCS map
where at each sky-position is associated an energy dispersion matrix.
The latter is a representation of the energy resolution as a function of the
true energy only repesented by the sub-class \code{EDispKernel}.

\subsubsection{Background model}
The backgropund rate can be represented in \gammapy as either a 2-dimensional
data structure (\code{Background2D}) of count rate normalised per steradians and energy at different
reconstructed energies and offset angles or as rate per steradians and energy, as a
function of reconstructed energy and detector coordinates (\code{Background3D}).
In the former, the background is expected to follow a radially symmetric shape,
while in the latter, it can be more complex.

@@

The models follow a naming scheme which contains the category as a suffix to
the class name. The spectral models include a special class of normed models,
named using the \code{NormSpectralModel} suffix.
These spectral models feature a dimension-less \textit{norm} parameter
instead of an \text{amplitude} parameter with physical units. They
must be used along with another spectral model, as a multiplicative correction
factor according to their spectral shape. They can be typically used for
adjusting template based models, or adding taking into account
the absorbtion effect on 
\gammaray spectra caused by the extra-galactic background light (EBL) (\code{EBLAbsorptionNormSpectralModel})
model. \gammapy supports a variety of EBL absorbtion models, such as
\cite{Franceschini2008}, \cite{Finke2010} and \cite{Dominguez2011}.

@@

The model gallery provides a visual overview of the available models in
\gammapy. Most of the analytic models  commonly used in \gammaray astronomy are
built-in. We also offer a wrapper to radiative models implemented in the Naima
package~\citep{naima}. The modeling framework can be easily extended with
user-defined models. For example agnpy models that describe leptonic radiative
processes in jetted Active Galactic Nuclei (AGN) can be wrapped into
\gammapy~\citep[see Section 3.5 of ][]{2021arXiv211214573N} .

@@

The Fit class provides a uniform interface to multiple fitting backends:

\begin{itemize}
	\item \iminuit~\citep{iminuit}
	\item \code{scipy.optimize}~\citep{2020SciPy-NMeth}
	\item Sherpa~\citep{sherpa-2011}
\end{itemize}

Note that, for now, covariance matrixand errors are computed only for the fitting with 
\iminuit. However depending on
the problem other optimizers can better perform, so sometimes it can be useful
to run a pre-fit with alternative optimization methods. In future we plan to
extend the supported fitting backends, including for example solutions based on Markov chain Monte Carlo.
\footnote{a prototype is available in gammapy-recipes,
	\url{https://gammapy.github.io/gammapy-recipes/_build/html/notebooks/mcmc-sampling-emcee/mcmc_sampling.html}
}

@@

Optionally it can also compute more advanced quantities
such as assymmetric flux errors, flux upper limits
and one dimensional profiles of the fit statistic,
which show how the likelihood functions varies with
the flux norm parameter around the fit minimum.
This information is useful for inspecting the quality
of a fit, for which asymptomatically a parabolic
shape of the profile is expected at the best fit
values.

@@

The code example shown in Figure~\ref{fig*:minted:gp_estimators} shows how to use
the \code{TSMapEstimator} objects with a given input \code{MapDataset}.
In addition to the model it allows to specify the energy
bins of the resulting flux and TS maps.

@@

\subsection{gammapy.analysis}
\label{ssec:gammapy-analysis}
The \code{gammapy.analysis} sub-module provides a high-level interface (HLI) for the most
common use cases identified in \gammaray analyses. The included classes and methods
 can be used in Python scripts, notebooks or as commands within \texttt{IPython}
sessions. The HLI can also be used to automatise
workflows driven by parameters declared in a configuration file in YAML format.
This way, a full analysis can be executed via a single command line taking the
configuration file as input.

The \code{Analysis} class has the responsibility of orchestrating of the workflow
defined in the configuration \code{AnalysisConfig} objects and triggering the execution of
the \code{AnalysisStep} classes that define the identified common use cases. These
steps include the following: observations selection with the \code{DataStore},  data
reduction, excess map computation, model fitting, flux points estimation, and
light curves production.

@@

\subsection{1D Analysis}
\label{ssec:1d-analysis}
One of the most common analysis cases in \gammaray astronomy is measuring the
spectrum of a source in a given region defined on the sky, in conventional
astronomy also called \textit{aperture photometry}. The spectrum is typically measured
in two steps: first a parametric spectral model is fitted to the data and
secondly flux points are computed in a pre-defined set of energy bins. The
result of such an analysis performed on three simulated CTA observations is
shown in Figure~\ref{fig:cta_galactic_center}. In this case the spectrum was
measured in a circular aperture centered on the Galactic Center, in
\gammaray~astronomy often called "on region". For such analysis the users first
chooses a region of interest and energy binning, both defined by a
\code{RegionGeom}. In a second step the events and the IRFs are binned
into maps of this geometry, by the \code{SpectrumDatasetMaker}. All the data and
reduced IRFs are bundled into a \code{SpectrumDataset}. To estimate
the expected background in the ON region a 'reflected regions' background
method was used~\cite{Berge07}, represented in \gammapy by the
\code{ReflectedRegionsBackgroundMaker} class. The resulting reflected regions are
illustrated for all three observations on top of the map of counts. After
reduction of the data it was modelled using a forward-folding method and a
power law spectral shape, using the \code{PowerLawSpectralModel} and \code{Fit} class.
Based on this best fit model the final flux points and corresponding
log-likelihood profiles are computed using the \code{FluxPointsEstimator}.

@@

The result of the joint fit is displayed in
Figure~\ref{fig:multi_instrument_analysis}. We remark that the objective of this
exercise is illustrative. We display the flexibility of \gammapy in
simultaneously fitting multi-instrument data even at different levels of
reduction, without aiming to provide a new measurement of the Crab Nebula
spectrum.

@@

\subsection{Broadband SED Modeling}
\label{ssec:broadband-sed-modeling}
By combining \gammapy with astrophysical modelling codes, users can also fit
astrophysical spectral models to \gammaray data. In \gammaray
astronomy one typically observes two radiation production
mechanisms, the so-called hadronic and leptonic scenarios.
There are several Python packages that are able to model
the \gammaray emission, given a physical scenario. Among those
packages are  Agnpy~\citep{agnpy}, Naima~\citep{naima}, Jetset~\citep{jetset}
and Gamera~\citep{gamera}.
Tyically those emission models predict broadband emission from
radio, up to the very high-energy \gammaray range.
By relying on the multiple dataset types in \gammapy those
data can be combined to constrain such a broadband emission model.
\gammapy provides a built-in \code{NaimaSpectralModel} that allows
users to wrap a given astrophysical emission model from the
naima package and fit it directly to \gammaray data.

As an example of this application, we use the same multi-instrument
dataset described in the previous section and we fit it with an inverse
Compton model computed with naima and wrapped in the \gammapy fitting
 routine with the NaimaSpectraModel class. We consider a log-parabolic
electron distribution scattering: the synchrotron radiation produced by
the same electrons; near and far infrared photon fields and the cosmic
microwave background (CMB). We adopted the prescription on the target
photon fields provided in the
documentation of the \textit{Naima} package \footnote{\url{https://naima.readthedocs.io/en/stable/examples.html\#crab-nebula-ssc-model}}.
The best-fit inverse Compton spectrum is represented with a red dashed line in
Figure~\ref{fig:multi_instrument_analysis}.

@@

Several automated tasks are set as \github
actions\footnote{\url{https://github.com/features/actions}}, blocking the
processes and alerting developers when fails occur. This is the case of the
continuous integration workflow, which monitors the execution of the test coverage
suite\footnote{\url{https://pytest.org}} using datasets from the
\textit{gammapy-data} repository\footnote{\url{https://github.com/gammapy/gammapy-data}}. Tests scan not only the codebase but also the
code snippets present in docstrings of the scripts and in the RST documentation
files, as well as in the tutorials provided in the form of Jupyter notebooks.

Other automated tasks, executing in the \textit{gammapy-benchmarks}\footnote{\url{https://github.com/gammapy/gammapy-benchmarks}} repository,
are responsible for numerical validation tests and benchmarks monitoring. Also
tasks related with the release process are partially automated, and every
contribution to the codebase repository triggers the documentation building and
publishing workflow within the \textit{gammapy-docs} repository\footnote{\url{https://github.com/gammapy/gammapy-docs}}
(see Sec.~\ref{ssec:software-distribution} and Sec.~\ref{ssec:documentation-and-user-support}).

@@

We have also set up a solution for users to share recipes as Jupyter notebooks
that do not fit in the \gammapy core documentation but which may be relevant as
specific use cases. Contributions happen via pull requests to the
\textit{gammapy-recipes} \github repository and merged after a short review. All
notebooks in the repository are tested and published in the \gammapy recipes
webpage\footnote{\url{https://gammapy.github.io/gammapy-recipes}} automatically
using \github actions.

A growing community of users is gathering around the Slack
messaging\footnote{\url{https://gammapy.slack.com}} and \github
discussions\footnote{\url{https://github.com/gammapy/gammapy/discussions}}
support forums, providing valuable feedback on the \gammapy functionalities,
interface and documentation. Other communication channels have been set like
mailing lists, a Twitter account\footnote{\url{https://twitter.com/gammapyST}},
regular public coding sprint meetings, hands-on session within collaborations,
weekly development meetings, etc.

@@

\section{Reproducibility}
\label{sec:reproducibility}
One of the most important goals of the \gammapy project is to support open and
reproducible science results. Thus we decided to write this manuscript
openly and publish the Latex source code along with the associated
Python scripts to create the figures
in an open repository~\footnote{\url{https://github.com/gammapy/gammapy-v1.0-paper}}.
This \github repository also documents the history of the creation
and evolution of the manuscript with time. To simplify the reproducibility
of this manuscript including figures and text, we relied on the tool
\textit{showyourwork}~\citep{Luger2021}. This tool coordinates the building
process and software as well as data dependencies such, that the complete
manuscript can be reproduced with a single \code{make} command, after
downloading the source repository. For this we provide
detailed instructions online\footnote{\url{https://github.com/gammapy/gammapy-v1.0-paper/blob/main/README.md}}.
Almost all figures in this manuscript provide a link
to a Python script, that was used to produce it. This means all
example analyses presented in Sec.\ref{sec:applications} link to
actually working Python source code.
